SAT Solving for Vesicle Traffic Systems in Cells

### 2017-06-22 paper/bio.tex
expmt
first cut

\begin{itemize}
\item \mukund{Detailed BIO motivation}
\item \srivas{Detailed BIO to Graph problem}
\end{itemize}

@@

\begin{itemize}
\item 3 connected
\item  ....
\end{itemize}

>>>>

\begin{itemize}
\item \mukund{Detailed BIO}
\item \srivas{Detailed BIO to Graph problem}
\end{itemize}

The whole cell should be in stable condition i.e flow of the molecule
is balanced at every point and globally. Each molecule leaving its
source node on a vesicle should come back in a cycle.

@@

\begin{itemize}
\item Vesicular traffic network.
\item Recycling of molecules, steady state.
\item Maximum connectivity [LGC, Least guarantee connectivity] and minimal connective [LRC (least required connectivity)] (rather than necc and suff cond) required by the graph. 
\item Connected-ness. 3 and 4.
\item Pairing mechanism (matrix).
\item Fusion rules, inhibition/regulation mechanism, Pairing inhibition.
\item Potential to connect.
\item Property of interest.
\item Reachability definition.
\end{itemize}

### 2017-06-22 paper/encoding.tex
first cut

In this section, we will formally define VTSs as a labelled graph
and the conditions on the graphs that are discussed in the previous
section.
Subsequently, we will present the SAT encoding of the search problems
for the graphs that satisfies the conditions.

@@

\begin{df}
  A VTS $G$ is tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the network, 
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes  
  \item $ \nodes \times (\powerset{\mols}-\emptyset) \times \nodes \subseteq \edges$ is the set of edges
    with molecule sets as labels
  \item $\pairs : \mols \times \mols$ is pairing relating
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges
  \end{itemize}
\end{df}
$\pairs$ defines which molecules can fuse with which molecules.
%
Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
$\nodef$ and $\edgef$ are used to define activity of molecules on
nodes and edges.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
\ashu{define missing conds V1-V7}

@@

Based upon earlier discussion, we need to answer the following search
question for ......\ashu{match with the earlier text}
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for stable and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols = \mu|$, and
$G$ is not $k$-connected.

@@

\item Boolean variable $p_{k,k'}$ indicates if $(m,m') \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}

We also have auxiliary variables that will help us encode the properties 
\begin{enumerate}
\item $a_{i,m}$ indicates that molecule $m$ is active at node $i$, i.e., $\nodef(m,L(i))$
  holds
\item $b_{i,j,q,m}$ indicates that molecule $m$ is active at $q$th edge $(i,M',j)$ between $i$ and $j$, i.e., $\edgef(m,M')$ holds
\item $r_{i,j,m,z}$ indicates that there is a path from $i$ to $j$ of length $z$
\item $d_{i,j,q}$ indicates $q$th edge between $i$ and $j$ is dropped(used in encoding
  $k$-connected property)
\end{enumerate}

% Every variable is a Boolean bit, for example, $n_{0,1}$ represents node 0
% with type 1 molecule present. Assignment to the Boolean variable
% n(0,1) to True will mean that at node 0 type 1 molecule is present.

% Basic notation for the formula we are trying to generate.
% \begin{enumerate}

% \item edge = e(i,j,q)      
%       we will assign e(i,j,q) for a qth edge between node i and j.

% \item dump = d(i,j,q)
%       to represent drop of edge(i,j,q) from the graph we will use d(i,j,q).

% \item presence node = n(i,k) 
% 	  kth molecule is present on the node i.

% \item active\_node = a(i,k) 
%       kth molecule is active on the node i.

% \item presence\_edge = e(i,j,q,k)
%       There is an qth edge between node i and j with kth molecule present.

% \item active\_edge = b(i,j,q,k)
%       There is an qth edge between node i and j with kth molecule active.

% \item reachability = r(i,j,k,z)
%       Node i and j are reachable with kth molecule present in z steps.

% \item pairing matrix = p(k,k')
%       Molecule k and k' are pairing molecules in the matrix.

% \item Type of the Boolean function = sorts (takes k-1 arguments Bool and return Bool) 
% \item Boolean function on nodes Function (fn\_{m},*sorts)

% \item Boolean function on edge Function(fe\_{m},*sorts)

% \end{enumerate}

@@

% \subsubsection{Constraints on the structure of the edges, nodes and }
\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints~\eqref{eq:f0} and~\eqref{eq:c3}
to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align}
  \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}\tag{V1}\label{eq:f0}
\end{align}

If a molecule is active on an edge, it should be present on the edge.
%
\begin{align}
  \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\tag{V2}\label{eq:f1}
\end{align}

@@

\subsubsection{Well-fused constraints}
Boolean constraint~\eqref{eq:fuse1} encodes that each edge must fuse with
destination node.
%
Boolean constraint~\eqref{eq:fuse2} encodes the each should not be able to
potentially fuse with any other node.

@@

% F4: For an edge to be valid, at least one SNARE pair on the vesicle and target compartment must be active and have a non-zero entry in the pairing matrix.  
\begin{align}
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies \bigvee_{m,m^{\prime}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \tag{V8}\label{eq:fuse1}  \\
\bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies \neg \bigvee_{j \neq j^{\prime}, m^{\prime\prime}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \tag{V9}\label{eq:fuse2}  
\end{align}

@@

\subsubsection{Regulation on nodes and edges}
%
Regulation of the molecules is based on the kind of the model under
consideration.
%
For example, model 1 (-V 1) has two different arbitrary
Boolean functions dictating the activity of the molecules on edge and
nodes.
%
For variation 3 (-V 3) activity on edge is driven by pairing
inhibition.

We use two Boolean variables A\_* to build these two constraints. \newline

A\_nn: No regulation on nodes i.e all the present molecules on the nodes are active.
\[
\bigwedge\limits_{i,m} n_{i,m} = a_{i,m} \,
\]

A\_nb: Boolean regulation on nodes. Activity of a molecule k on the node is defined as a Boolean function of presence of other molecule present on that node.
\[ \bigwedge\limits_{i,k} n_{i,m} \limplies a_{i,m} =   f_m (n_{i,1},\dots,n_{i,\mu}) \, \]

A\_en: No regulation on edges i.e all the present molecules on the edges are active.
\[ \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m} \, \]

A\_eb: Boolean regulation on edges. Activity of a molecule k on the node is defined as a Boolean function of presence of other molecule present on that node.
\[ \bigwedge\limits_{i,j,q,k} e_{i,j,q,k} \limplies b_{i,j,q,k} == f_e[k](\bigvee_{k' != k} e_{i,j,q,k'}) \, \]

A\_ep: Pairing inhibition. The activity of the molecules is defined by inhibition by other molecules based on the pairing matrix. 
\[ \bigwedge\limits_{i,j,q, k}  (e_{i,j,q,k} \land [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies \neg b_{i,j,q,k}) \, \]

\[ \bigwedge\limits_{i,j, q, k}  (e_{i,j,q,k} \land \neg [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies b_{i,j,q,k}) \, \]

@@

\subsubsection{Constraints for stability condition}
The whole cell should be in stable condition i.e flow of the molecule is balanced at every point and globally. Each molecule leaving its source node on a vesicle should come back in a cycle. Boolean variable F3 and F2 are used to model these constraints.

\textbf {Reachability definition and stability condition.}
We have encoded stable condition by using reachability definition. \newline

F3: Stability condition. For every leaving molecule, the source node is reachable from the target node with that molecule present, in maximum p steps.

% I have implemented: Changed the implementation in code too.
%\[ \bigwedge\limits_{i,j,q,k} e_{i,j,q,k} \limplies (r_{j,i,k,0}  \lor r_{j,i,k,1} \lor ... \lor r_{j,i,k,p}) \, \]

% Update: Better 
\[ \bigwedge\limits_{i,j,k} (\bigvee_{q} e_{i,j,q,k}) \limplies (r_{j,i,k,0}  \lor r_{j,i,k,1} \lor ... \lor r_{j,i,k,p}) \, \]

F2: Reachability definition. Node j is reachable from node i with kth molecule present in maximum p steps if either there is a direct edge between i-j with that molecule present or there is a direct edge between $i^{\prime \prime}$ $(i != i^{\prime \prime})$ and j with kth molecule present on that edge and $i^{\prime \prime}$ are reachable from i in p steps.

\[ \bigwedge\limits_{i,j,k,p} r_{i,j,k,p} \limplies (\bigvee_{q} \, e_{i,j,q,k} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,k}) \land r_{i^{\prime},j,k,p-1} ) \, \]

\subsubsection{Graph connectivity constraints}
To check the whether that n connected is a necessary condition, we remove (drop) n edges from the graph and if it disconnects the graph and we get an assignment we have an n connected satisfying graph. We can go down or up using -C \_ option.  \newline

D0: Only present edges can be dropped.
\[ \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \]

D1,D2: We are dropping c edges from the graph. exactly c are dropped.
\[ \sum_{i,j,q} d_{i,j,q} == c \]

D3:  Graph becomes disconnected. Ensure that there is no path between some nodes i,j in the underlying undirected graph. Different reachability.
\[ \bigwedge\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})]  \]

D4: New reachability definition for graph connectedness: dReachable. Node i,j are reachable either if there is a direct edge and its not dropped or there is an node $i^{\prime}$ such that, there is a direct edge between i, $i^{\prime}$ which is not dropped and $i^{\prime}$ and j is dReachable.

\[ \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' != i}  r^{\prime}(i',j) \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}(i,j)  \]

\subsection{Creating a solver and solving these constraints}
We will create a solver and add these build formula using these constraints to that and then check whether there a satisfiable assignment that satisfies all these constraints.

\begin{figure}[ht]
%\vspace{-3mm}
\begin{verbatim}
# Create a solver
s = Solver()
# Add these formulas to it.
s.add(A0, A1, C0, C1, C2, C4, C5, F0, F1, F2, F3, F4, F5, D0, D1, D2, D3)

# Check for the satisfiability assignment.
print s.check()   
\end{verbatim}
\caption{Solving for these constraints.}
\label{code:motivate}
\end{figure}

>>>>

In this section, we will formally define VTSs as a labelled graph
and the conditions on the graphs that are discussed in the previous
section.
%
Subsequently, we will present the SAT encoding of the search problems
for the graphs that satisfies the conditions.

@@

\begin{df}
  A VTS $G$ is tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the network, 
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes  
  \item $ \nodes \times (\powerset{\mols}-\emptyset) \times \nodes \subseteq \edges$ is the set of edges
    with molecule sets as labels
  \item $\pairs : \mols \times \mols$ is pairing relating
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges
  \end{itemize}
\end{df}
$\pairs$ defines which molecules can fuse with which molecules.
%
Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
$\nodef$ and $\edgef$ are used to define activity of molecules on
nodes and edges.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We call $G$ {\em well-structured} if 
Molecules $M$ is divided into two partitions $M_1$ and $M_2$
such that $M_1 \times M_2 \intersection P = \emptyset$, and
for each $(n,M',n')$, $n \neq n'$, 
$M' \subseteq \nlabel(n)$, and $M' \subseteq \nlabel(n')$.
%

@@

Based upon earlier discussion, we need to answer the following search
question for ......\ashu{match with the earlier text}
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for well-structured, stable, and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols = \mu|$, and
$G$ is not $k$-connected.

@@

\item Boolean variable $p_{k,k'}$ indicates if $(m,m') \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}
We also have auxiliary Boolean variables that will help us encode the well-fused and stability properties 
\begin{enumerate}
\item $a_{i,m}$ indicates that molecule $m$ is active at node $i$, i.e., $\nodef(m,L(i))$
  holds
\item $b_{i,j,q,m}$ indicates that molecule $m$ is active at $q$th edge $(i,M',j)$ between $i$ and $j$, i.e., $\edgef(m,M')$ holds
\item $r_{i,j,m,p}$ indicates that there is a path from $i$ to $j$ of length less than or equal to $p$
\end{enumerate}
For $k$-connected property, we also use the following auxiliary Boolean variables
\begin{enumerate}
\item $d_{i,j,q}$ indicates $q$th edge between $i$ and $j$ is dropped
\item $r'_{i,j}$ indicates that there is a path from $i$ to $j$ is a modified VTS
\end{enumerate}

@@

% \subsubsection{Constraints on the structure of the edges, nodes and }
\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints~\eqref{eq:f0} and~\eqref{eq:c3}
to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align}
  \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}\tag{V1}\label{eq:f0}
\end{align}
If a molecule is active on an edge, it should be present on the edge.
%
\begin{align}
  \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\tag{V2}\label{eq:f1}
\end{align}

@@

\subsubsection{Well-fused constraints}
Boolean constraint~\eqref{eq:fuse1} encodes that each edge must fuse with
its destination node.
%
Boolean constraint~\eqref{eq:fuse2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align}
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies \bigvee_{m,m^{\prime}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \tag{V8}\label{eq:fuse1}  \\
\bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies \neg \bigvee_{j \neq j^{\prime}, m^{\prime\prime}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \tag{V9}\label{eq:fuse2}  
\end{align}

@@

% F4: For an edge to be valid, at least one SNARE pair on the vesicle and target compartment must be active and have a non-zero entry in the pairing matrix.

@@

\subsubsection{Regulation on nodes and edges}
We are considering several models that differ in constraints on
the activity of molecules.
%
We will present~\eqref{eq:ann}-~\eqref{eq:aep} that encodes
the varying constraints.
%
The following formula encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} = a_{i,m}    \tag{Ann}\label{eq:ann}
\end{align}
The following formula encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align}
\bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\tag{Anb}\label{eq:anb}
\end{align}
The following formula encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\tag{Aen}\label{eq:aen}
\end{align}
The following formula encodes that activity of a molecule $m$ on the edge is
defined by a Boolean function $\edgef_m$ of presence of molecules present on that edge.
\begin{align}
   \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
  \tag{Aeb}\label{eq:aeb}
\end{align}
%
The following formula encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align}
   \bigwedge\limits_{i,j,q, m}  e_{i,j,q,m} \limplies(  [ ( \bigvee_{m'} p_{m,m'} \land \bigwedge_{m' != m} p_{m,m'} \limplies e_{i,j,q,m'})] \lequiv \neg b_{i,j,q,m})
  \tag{Aep}\label{eq:aep}
\end{align}
% \[ \bigwedge\limits_{i,j,q, k}  (e_{i,j,q,k} \land [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies \neg b_{i,j,q,k}) \, \]

% \[ \bigwedge\limits_{i,j, q, k}  (e_{i,j,q,k} \land \neg [ ( \bigvee_{k'} p_{k,k'} \land \bigwedge_{k' != k} p_{k,k'} \limplies e_{i,j,q,k'})] ) \limplies b_{i,j,q,k}) \, \]

@@

\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following formula recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i != i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align}
  \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
  \tag{R1}\label{eq:reach1}
\end{align}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.\ashu{disjunction needed??}
\begin{align}
 \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies (r_{j,i,m,0}  \lor r_{j,i,m,1} \lor ... \lor r_{j,i,m,p})  
  \tag{R2}\label{eq:reach2}
\end{align}

% Boolean variable
% F3 and F2 are used to model these constraints.

% % \textbf {Reachability definition and stability condition.}
% We have encoded stable condition by using reachability definition. \newline

% F3: Stability condition. For every leaving molecule, the source node is reachable from the target node with that molecule present, in maximum p steps.

% F2: Reachability definition. Node j is reachable from node i with kth molecule present in maximum p steps if either there is a direct edge between i-j with that molecule present or there is a direct edge between $i^{\prime \prime}$ $(i != i^{\prime \prime})$ and j with kth molecule present on that edge and $i^{\prime \prime}$ are reachable from i in p steps.

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
The following formulas encode the relevant constraints for reachability in the modified VTS.

The following formulas encode that only existing edges can be dropped
and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following formula, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' != i}  r^{\prime}(i',j) \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}(i,j)  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we are looking for a graph that is not $k$-connected,
the following formula encodes that there are nodes $i$ and $j$ such that
there is no path between them in the underlying undirected graph.
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

The key improvement in our work over the earlier tool is the
encoding of reachability, which was done using enumeration of paths.
%
We have encoded reachability in two different ways in~\eqref{eq:reach1} and~\eqref{eq:drop3}.
%
The reachability is recursively defined~\eqref{eq:drop3} and has trivial solution by
making all $r'$s true.
%
However, the trivial solution is disallowed by~\eqref{eq:drop3} and we find only the
solutions that captures the evidence of unreachability.
%
In~\eqref{eq:reach1}, we have added length of paths in our reachability encoding.
%
This is because~\eqref{eq:reach2} has only positive occurrences of the reachability variables
and circular dependencies in the recursive definition in~\eqref{eq:reach1} 
may result in spurious satisfying assignments that do not encode reachability.
%
By adding the path length, we break the circular dependencies and the formula remain polynomial in size.

% We can go down or up using -C \_ option.  \newline

\subsection{SAT problems different variants}
\ankit{Bring model variant table here}

% \subsection{Creating a solver and solving these constraints}
% We will create a solver and add these build formula using these constraints to that and then check whether there a satisfiable assignment that satisfies all these constraints.

% \begin{figure}[ht]
% %\vspace{-3mm}
% \begin{verbatim}
% # Create a solver
% s = Solver()
% # Add these formulas to it.
% s.add(A0, A1, C0, C1, C2, C4, C5, F0, F1, F2, F3, F4, F5, D0, D1, D2, D3)

% # Check for the satisfiability assignment.
% print s.check()   
% \end{verbatim}
% \caption{Solving for these constraints.}
% \label{code:motivate}
% \end{figure}
